I Compensate, therefore I Am \\ (accurate for DNN verification)

### 2024-01-18 Paper/Paper_noProof.tex
TYPO

For DNNs easy to verify, $\beta$-CROWN and PRIMA are excellent solutions, both very fast and very accurate, usually matching or getting very close to the upper bound on the $\%$ of verified images \cite{crown}. Compensation strength is not required in these contexts, and likely not useful, reason why we did not consider it.

We only considered DNNs with ReLU activation functions in this paper. The reason is that MILP, which is the basic block of Compensate($K$), and BaB (the basic block of $\beta$-Crown) can handle the ReLU activation function efficiently, and more complex activation functions are abstracted in other ways, which PRIMA handles (unlike $\beta$-CROWN). Still, the compensation paradigm is totally orthogonal to ReLU activation function. It is very likely that studying the compensation strength would be useful in this  context as well. We leave that for future work.

While accuracy of compensate($K$) can be among the highest of the efficient verifiers, runtime, while never blowing-up, can be on the slow side, in particular for larger DNNs. There are many optimizations possible. The simplest is to run $\beta$-CROWN first in the pipeline, and only use Compensate($K$) for those images for which $\beta$-CROWN cannot conclude, possibly using more core to compute more nodes in parallel. Another simple way is to modify the refined BaB of $\beta$-CROWN to run compensate($K$) on more layers than what exact MILP can be run, and finish the study using BaB. More involved would be to modify BaB to branch on unstable ReLU nodes guided by the compensation strength, which we experimentally proved to be an efficient trade-off. Also, an analysis of important nodes for the uncertain output nodes could be performed, allowing to compute much faster less accurate bounds for those nodes not important. Last, one could perform a Refinement Abstraction framework similar to \cite{atva}, \cite{elboher}, or \cite{SRGR}.

@@

In this paper, we presented the notion of compensating pairs of paths, with rationale why such a phenomenon creates inaccuracies hard to handle when verifying DNNs. We proved that this phenomenon is actually explaining entirely the inaccuracies, as in their absence, even very simple arithmetic algebra suffices to verify accurately DNNs. We verified that DNNs harder to verify indeed exhibits more paths with larger compensating strength. 
Based on this idea of compensating path, we proposed a novel efficient algorithm considering a subset $Z$ of the set of unstable ReLU nodes, that we proved to be fully accurate if all the compensating paths are covered. We experimentally tested that choosing $Z$ covering the paths with the strongest compensation result to very good accuracy, and runtime which can be favourable in some (but not all) cases. 
This proves that compensating strength is a promising novel concept which should be further studied to 
improve DNN verification. We finally proposed several ways to use the compensating strength to optimize the tools even more than they are, and leave that as future work.

>>>>

For DNNs easy to verify, $\alpha$-$\beta$-CROWN and PRIMA are excellent solutions, both very fast and very accurate, usually matching or getting very close to the upper bound on the $\%$ of verified images \cite{crown}. Compensation strength is not required in these contexts, and likely not useful, reason why we did not consider it.

We only considered DNNs with ReLU activation functions in this paper. The reason is that MILP, which is the basic block of Compensate($K$) (and BaB for $\beta$-Crown) can handle the ReLU activation function efficiently. Other activation functions are abstracted in more complex ways, e.g. by PRIMA (unlike $\beta$-CROWN). Still, the compensation paradigm is  orthogonal to the activation function. We believe that studying the compensation strength would be useful with other activation functions as well. We leave that for future work.

Accuracy of Compensate($K$) can be among the highest of the efficient verifiers.
Runtime, while never blowing-up, can be on the slow side though, in particular for larger DNNs. There are many optimizations possible. The simplest is to run $\alpha$-$\beta$-CROWN first in the pipeline, and only use Compensate($K$) for those images for which $\alpha$-$\beta$-CROWN cannot conclude, possibly using more CPU cores to compute more nodes in parallel. Another simple way is to modify the refined BaB of $\beta$-CROWN to run Compensate($K$) on more layers than what exact MILP can do, and finish the verification using BaB. More involved would be to modify BaB to branch on unstable ReLU nodes guided by the compensation strength, which we experimentally proved to be an efficient trade-off. Also, an analysis of important nodes for the uncertain output nodes could be performed, allowing to compute much faster less accurate bounds for those nodes not important. Last, one could perform a Refinement Abstraction framework similar to \cite{atva}, \cite{elboher}, or \cite{SRGR}.

@@

In this paper, we presented the notion of compensating pairs of paths, with rationale why such a phenomenon creates inaccuracies hard to handle when verifying DNNs. We proved that this phenomenon is actually explaining entirely the inaccuracies, as in their absence, even very simple arithmetic algebra suffices to verify accurately DNNs. We verified that DNNs harder to verify indeed exhibits more paths with larger compensating strength.

Based on this idea of compensating path, we proposed a novel efficient algorithm considering a subset $Z$ of the set of unstable ReLU nodes, that we proved to be fully accurate if all the compensating paths are covered. We experimentally tested that choosing $Z$ covering the paths with the strongest compensation result to very good accuracy, and runtime which can be favourable in some (but not all) cases. 
This proves that compensating strength is a promising novel concept which should be further studied to 
improve DNN verification. We finally proposed several ways to use the compensating strength to optimize the tools even more than they are, and leave that as future work.
